Reducing RES Droughts through the integration of wind and solar PV

### 2025-04-01 Article/RES_droughts_Morin_ReEn_R2.tex
Update Reviewed Version of the Article

Alternative methods adjust the CF threshold dynamically over the year to account for seasonal variations in renewable production. Raynaud et al.~\citep{raynaud2018drought} defined RES droughts as sequences of days with renewable electricity generation below a threshold that varies seasonally, a methodology later adapted for India~\citep{gangopadhyay2022drought}. Building on this, Kapica et al.~\citep{kapica2024drought} compared the likelihood of increased RES droughts in Europe under different climate models. Other studies have defined RES droughts based on deviations from daily mean production: Rinaldi et al.~\citep{rinaldi2021drought} applied these in the U.S. Western Interconnection to quantify the benefits of long-term storage, while Brown et al.~\citep{brown2021drought} examined weekly timescales to explore meteorological influences on the most severe RES drought events. Another method defines RES drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}. This approach identifies periods of unusually low generation relative to historical production levels, using the lowest production percentiles. Bracken at al.~\citep{bracken2024drought} used this approach to analyse RES droughts at different time scales in the U.S.~\citep{bracken2024drought}, and Lei et al.~\citep{lei2024drought} used it to quantify RES droughts in wind-PV-hydro systems in China.

@@

In this paper, the focus is exclusively on renewable electricity generation, which allows us to maintain physical models that do not consider the behavioural influence of demand. A fixed threshold approach is used to define RES droughts, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities. The case study used in this paper is Ireland, a region with a strong reliance on wind power and ambitious targets for solar PV power expansion. This provides valuable insights into the potential benefits of diversifying the share of wind and solar PV on RES droughts.

>>>>

Alternative methods adjust the CF threshold dynamically over the year to account for seasonal variations in renewable production. Raynaud et al.~\citep{raynaud2018drought} defined RES droughts as sequences of days with renewable electricity generation below a threshold that varies seasonally, a methodology later adapted for India~\citep{gangopadhyay2022drought}. Building on this, Kapica et al.~\citep{kapica2024drought} compared the likelihood of increased RES droughts in Europe under different climate models. Other studies have defined RES droughts based on deviations from daily mean production: Rinaldi et al.~\citep{rinaldi2021drought} applied these in the U.S. Western Interconnection to quantify the benefits of long-term storage, while Brown et al.~\citep{brown2021drought} examined weekly timescales to explore meteorological influences on the most severe RES drought events. Another method defines RES drought indices based on metrics commonly used in hydro-meteorology to characterise RES droughts~\citep{allen2023drought}. This index was also applied to the U.S. by Bracken et al.~\citep{bracken2024drought}, revealing a consistent increase in the RES drought magnitude when load is considered, despite showing differing results across regions. The consideration of demand introduces a completely different factor to generation: weather is not the only driver, as the societal component plays a large role.

@@

In this paper, the focus is exclusively on renewable electricity generation, which allows us to maintain physical models that do not consider the behavioural influence of demand. A fixed threshold approach is used to define RES droughts, which facilitates consistent inter-comparison between scenarios with different installed wind and solar PV capacities.  The case study used in this paper is Ireland, a region where most RES generation comes from wind power and with ambitious targets for solar PV power expansion. This provides valuable insights into the potential benefits of adding solar PV installations in wind-dominated countries.
